% Add the following subsections
% - Problem -> Motivation
% - Solution -> Summary of Solutions
% - Detailed description of the solution
% - Architecture and Design
% - Examples
% - Comparison with the other solutions, state of the art

\input{papers/template/style.tex}

\graphicspath{{papers/Causify_AIgentic_Development_System/}}

\begin{document}
  \title{Causify AIgentic Development System}
  \maketitle

  % ############################################################################
  \section{Introduction}
  Software development workflows are becoming more complex as they adapt
  to the demands of large-scale systems and modern collaborative
  development practices. As teams and codebases grow, companies face the
  challenge of organizing both effectively. When it comes to structuring
  the codebase, two main approaches emerge: monorepos and multi-repos.
  Monorepos consolidate all code into a single repository, simplifying version
  control but carrying a risk of scalability and maintainability issues.
  Conversely, multi-repos store the code in logically separated repositories,
  easier to manage and deploy but more difficult to keep in sync.

  In this paper, we propose Causify dev system, an alternative hybrid
  solution: a modular system architecture built around runnable directories.
  Although independent, these directories maintain cohesion through shared
  tooling and environments, offering a straightforward and scalable way
  to organize the codebase while ensuring reliability in development, testing,
  and deployment.

  % ############################################################################
  \section{Contributions}

  This paper presents the Causify development system, a novel approach
  to software development infrastructure that addresses the limitations
  of traditional monorepo and multi-repo architectures, development vs deployment,

  The main contributions are:

  \begin{itemize}
    \item \textbf{Human / AI development system}:
      A development system designed to enable AI agentic coding, to increase
      the efficiency of human developers, and to allow AI coding agents and
      humans to collaborate.

    \item \textbf{Runnable directories architecture}: A hybrid organizational
      model that combines the modularity of multi-repos with the environment
      consistency of monorepos, enabling independent development
      lifecycles while maintaining unified tooling and testing
      capabilities.

    \item \textbf{Containerized development workflows}: 
      A Docker-based approach to development that ensures reproducible
      environments across development, testing, and production stages.

    \item \textbf{Dockerized Executables}: A containerization pattern for
      individual development utilities that packages tools with their
      dependencies in slim Docker images, enabling execution without local
      installation.
  \end{itemize}

  This paper is organized as in the following
  TODO(gp)

  % ############################################################################
  \input{papers/Causify_AIgentic_Development_System/part1.code_quality.tex}

  % ############################################################################
  \input{papers/Causify_AIgentic_Development_System/part2.runnable_directory.tex}

  % ############################################################################
  \input{papers/Causify_AIgentic_Development_System/part3.containerized_development.tex}

  % ############################################################################
  \input{papers/Causify_AIgentic_Development_System/part4.AI_optimized_development.tex}

  %- [1]
  %  [Mono vs. multi-repo](https://free.gitkraken.com/hubfs/Mono_v_Multi-Repo_debate_2023.pdf)
  %- [2]
  %  [Why Google stores billions of lines of code in a single repository](https://dl.acm.org/doi/10.1145/2854146)
  %- [3]
  %  [What it is like to work in Meta's (Facebook's) monorepo](https://blog.3d-logic.com/2024/09/02/what-it-is-like-to-work-in-metas-facebooks-monorepo/)
  %- [4]
  %  [Microsoft: How "Mono-repo" and "One Infra" Help Us Deliver a Better Developer Experience](https://devblogs.microsoft.com/appcenter/how-mono-repo-and-one-infra-help-us-deliver-a-better-developer-experience/)
  %- [5]
  %  [Uber: Faster Together: Uber Engineering's iOS Monorepo](https://www.uber.com/blog/ios-monorepo/)
\end{document}
