% ##############################################################################
  \section{Policy Model: From Tenants to Capabilities}
  This section specifies the \policy{} model that powers \uiPolicy{}. The
  key design choice is to store tenant experience as data and \emph{materialize
  it server-side} into a ``UI contract'' that the web tier can consume
  without inventing permissions in the browser.

% ==============================================================================
  \subsection{Identity claims and role derivation}
  Authentication is delegated to Amazon Cognito user pools
  \cite{aws_cognito_user_pools}. The control plane receives an ID token and
  derives two primitives:

  \begin{itemize}
    \item \textbf{Actor tenant:} extracted from a custom JWT claim (e.g.,
      \texttt{custom:tenantId}).

    \item \textbf{Actor role:} derived from \texttt{cognito:groups} \cite{aws_cognito_groups}.
  \end{itemize}

  Groups follow a simple convention:
  \begin{center}
    \texttt{tenant\_id:role} \quad (e.g., \texttt{acme:member}, \texttt{causify:admin})
  \end{center}

  Users may belong to multiple groups for a tenant. We resolve role by precedence:
  \begin{center}
    \texttt{admin} $>$ \texttt{member} $>$ \texttt{demo}
  \end{center}

  This rule is intentionally boring. Boring is good. Boring prevents the
  ``surprise admin'' incident that ruins everyone's weekend.

  \begin{algorithm}
    [H]
    \caption{Role resolution from Cognito groups (simplified)}
    \label{alg:role_resolution}
    \begin{algorithmic}
      [1] \Require tenant\_id, groups \If{groups is empty} \Return \texttt{None}
      \EndIf \State valid $\gets$ \{ g $\in$ groups : \texttt{":"} $\in g$
      and g startsWith \texttt{tenant\_id:} \} \If{valid is empty} \Return
      \texttt{None} \EndIf \State roles $\gets$ extract suffix after
      \texttt{":"} for each g in valid \State order $\gets$ [\texttt{admin},
      \texttt{member}, \texttt{demo}] \State \Return argmin\_r index(r
      in order) (default to end if unknown)
    \end{algorithmic}
  \end{algorithm}

% ==============================================================================
  \subsection{Impersonation policy: who may \viewas{} whom}
  \uiPolicy{} includes a first-class \viewas{} mechanism to support
  operations, QA, and demo flows without shared admin credentials. This is
  also the fastest way to reproduce tenant-specific UI bugs: you \emph{become}
  the bug (politely).

  We compute an explicit \texttt{impersonatable} allow-list in the
  control plane:

  \begin{itemize}
    \item \textbf{Super admin:} a designated tenant/role pair may
      impersonate all active tenants and roles (enumerated from the tenant
      registry table).

    \item \textbf{Tenant admin:} an admin may impersonate all roles
      \emph{within} their tenant.

    \item \textbf{Everyone else:} may impersonate only what their group
      memberships allow (e.g., if they are in \texttt{acme:demo} and
      \texttt{acme:member}, they may \viewas{} either).
  \end{itemize}

  This allow-list is returned to the web tier, enabling the UI to offer
  safe tenant/role selection while keeping enforcement server-side.

% ==============================================================================
  \subsection{Tenant policy as data (DynamoDB)}
  Tenant variability is expressed as \policy{}: a DynamoDB record keyed
  by \texttt{tenant\_id}. The record contains tenant metadata plus a
  role map under \texttt{group}. Each role entry provides:

  \begin{itemize}
    \item \texttt{modules}: the enabled module identifiers for that role.

    \item \texttt{moduleLayout}: a map of JSON \emph{strings} that
      represent UI layout/config fragments. These fragments are parsed server-side
      during materialization.
  \end{itemize}

  \begin{samepage}
  \begin{lstlisting}[style=json,caption={Illustrative tenant policy record (close to production shape). Role entries live under \texttt{group}. Layout fragments are stored as JSON strings and parsed server-side during materialization into structured objects.},label={lst:tenantpolicy_v2}]
{
  "tenant_id": "acme",
  "domain": "acme.example",
  "status": "active",
  "allowed": true,
  "schema_prefix": "acme",
  "group": {
    "demo": {
      "modules": ["horizon"],
      "moduleLayout": {
        "common": "{ \"sideBar\": { \"defaultCollapsed\": true } }"
      }
    },
    "member": {
      "modules": ["horizon", "grid"],
      "moduleLayout": {
        "common": "{ \"sideBar\": { \"defaultCollapsed\": false }, \"feedback\": { \"enabled\": true } }",
        "grid":   "{ \"sidebar\": { \"enabled\": true }, \"table\": { \"pageSize\": 50 } }"
      }
    },
    "admin": {
      "modules": ["horizon", "grid", "datamap"],
      "moduleLayout": {
        "common": "{ \"settings\": { \"enabled\": true }, \"sideBar\": { \"showAdmin\": true } }"
      }
    }
  }
}
\end{lstlisting}
  \end{samepage}

  \paragraph{Validation and failure behavior.}
  Because this policy is runtime-loaded, the backend must treat it as
  untrusted input. We apply strict parsing: if a \texttt{moduleLayout} fragment
  fails JSON parsing, the control plane rejects the request with a
  server error. This avoids silently serving half-valid policies that produce
  haunted UIs (buttons moving on their own, etc.).

% ==============================================================================
  \subsection{Materialization: the \texttt{/v2/me} UI contract}
  The control plane returns a \emph{materialized} profile via \texttt{/v2/me}.
  This response is the UI contract: it tells the web tier who the actor is,
  what effective tenant/role is in force (via \viewas{}), what modules
  exist, and what layout fragments should shape the interface.

  Crucially, \texttt{/v2/me} returns both:

  \begin{itemize}
    \item \textbf{\texttt{actor}} --- immutable identity (who is authenticated),

    \item \textbf{\texttt{view\_as}} --- effective context (who the actor
      is operating \emph{as}).
  \end{itemize}

  \begin{samepage}
  \begin{lstlisting}[style=json,caption={Simplified \texttt{/v2/me} response. The \texttt{view\_as} block materializes modules + layout for the effective tenant/role.},label={lst:v2me}]
{
  "actor": {
    "name": "Ada Lovelace",
    "email": "ada@acme.example",
    "tenant_id": "causify",
    "user_role": "admin",
    "groups": ["causify:admin", "acme:member"]
  },
  "view_as": {
    "view_as_self": false,
    "tenant_id": "acme",
    "user_role": "member",
    "modules": ["horizon", "grid"],
    "module_layout": {
      "common": {
        "sideBar": { "defaultCollapsed": false, "showAdmin": true },
        "feedback": { "enabled": true },
        "topBar": { "enabled": true, "title": "Horizon" }
      },
      "horizon": {
        "enabled": true,
        "pageTitle": "Inventory Management",
        "dashboard": { "enabled": true },
        "gridView": { "sidePanel": { "enabled": true } }
      }
    }
  },
  "impersonatable": [
    { "tenant_id": "acme", "roles": ["demo", "member", "admin"] },
    { "tenant_id": "xyz", "roles": ["member"] }
  ]
}
\end{lstlisting}
  \end{samepage}

  \paragraph{Caching.}
  The control plane supports standard caching semantics (ETag-based conditional
  requests) and short TTL caching headers. This keeps the UI responsive while
  still enabling policy edits to take effect quickly.

  \paragraph{Why this matters.}
  Once \texttt{/v2/me} exists, the UI no longer needs to infer the user's
  world. It can simply render the world it was handed.
