% ##############################################################################
  \section{Safe Impersonation (\viewas): Support Without Chaos}
  Impersonation is one of those features that operators love and security
  teams fear. OWASP explicitly calls out account impersonation risks
  because it can collapse accountability and enable privilege escalation
  \cite{owasp_impersonation}. We therefore design \viewas{} as a
  controlled capability with explicit guardrails.

% ==============================================================================
  \subsection{Actor vs.\ effective context}
  Every request has:
  \begin{itemize}
    \item \textbf{actor}: the authenticated user identity (tenant, role,
      groups),

    \item \textbf{\viewas}: an optional effective tenant/role used for policy
      resolution (modules + layout) and downstream scoping.
  \end{itemize}

  The system returns \texttt{view\_as.view\_as\_self} so the UI can signal
  when the user is impersonating. This is not cosmetic; it prevents support
  engineers from accidentally doing ``real'' operations in the wrong workspace.

% ==============================================================================
  \subsection{Protocol: headers, signing, and session storage}
  Impersonation is conveyed via a semicolon-delimited header:
  \begin{center}
    \texttt{X-View-As: tenant\_id=<id>;role=<role>}
  \end{center}

  The web tier generates and stores these headers server-side (session) and
  attaches them to control-plane calls. To add defense-in-depth, the web
  tier also generates an HMAC signature:
  \begin{center}
    \texttt{X-View-As-Sig: HMAC\_SHA256(secret, X-View-As)}
  \end{center}

  The important operational detail is that the signed headers are stored
  on the server side of the web tier session. The browser does not hold
  the secret and does not directly manage impersonation headers.

  \paragraph{Note.}
  The control plane currently validates \viewas{} using the allow-list; verifying
  the signature at the API tier is recommended and listed as future work
  (Section~\ref{sec:futurework}).

% ==============================================================================
  \subsection{Enforcement: allow-lists, not vibes}
  The control plane rejects impersonation unless:
  \begin{itemize}
    \item \texttt{tenant\_id} exists in \texttt{impersonatable},

    \item \texttt{role} is valid for that tenant entry,

    \item required header keys are present.
  \end{itemize}

  This prevents users from ``guessing'' a tenant ID and role combination
  and hoping something accepts it. It also prevents a web tier bug from
  becoming a permission escalation, because the control plane is the final
  gate.

% ==============================================================================
  \subsection{Downstream scoping: tenant-aware context}
  Once validated, \viewas{} affects:
  \begin{itemize}
    \item \textbf{policy resolution:} modules and layout are fetched for
      effective tenant/role,

    \item \textbf{tenant scoping:} the system resolves a tenant-specific
      prefix (e.g., schema prefix) from the effective tenant record to
      scope downstream services and data access.
  \end{itemize}

  This last part is easy to overlook, but it is where impersonation
  becomes dangerous if implemented casually: \viewas{} must not only
  change what the UI shows, it must correctly scope what the backend \emph{touches}.

% ==============================================================================
  \subsection{Operational UX: visible, reversible, auditable}
  A safe \viewas{} workflow is more than a header:
  \begin{itemize}
    \item \textbf{Visible:} the UI clearly indicates the effective
      tenant/role.

    \item \textbf{Reversible:} users can return to self with a single
      action.

    \item \textbf{Auditable:} logs and traces attribute actions to the
      actor even when operating under \viewas{}.
  \end{itemize}

  The principle is simple: \viewas{} is a support tool, not a second
  identity. The actor remains accountable at all times.

% ==============================================================================
  \subsection{Recommended extensions}
  We found two extensions particularly useful:
  \begin{itemize}
    \item \textbf{Read-only mode:} allow \texttt{X-View-As} to carry an explicit
      \texttt{mode=read-only} hint (and enforce it server-side) to
      prevent accidental state changes during support sessions.

    \item \textbf{Signature verification at the API tier:} validate
      \texttt{X-View-As-Sig} in the control plane for defense-in-depth.
  \end{itemize}
