% ##############################################################################
  \section{Implementation Notes}
  This section grounds \uiPolicy{} in concrete engineering choices. The goal
  is not to sell a specific stack, but to show the mechanics that make
  the UI \emph{provably less prone to lying}.

% ==============================================================================
  \subsection{Control plane: Lambda as policy materializer}
  The control plane is an AWS Lambda REST API built with Powertools for AWS
  Lambda (Python) \cite{aws_powertools}. The handler uses an \texttt{APIGatewayRestResolver}
  and exposes both \texttt{/v1/me} (legacy) and \texttt{/v2/me} (actor/\viewas{}
  model).

  Three middlewares do most of the heavy lifting:

  \begin{enumerate}
    \item \textbf{Authentication \& context middleware.} It validates
      the JWT, extracts \texttt{custom:tenantId}, derives the role from \texttt{cognito:groups},
      computes \texttt{available\_impersonations}, validates \texttt{X-View-As}
      (if present) against that allow-list, and resolves \texttt{schema\_prefix}
      from the effective tenant.

    \item \textbf{ETag middleware.} It supports conditional requests via
      \texttt{If-None-Match} and returns \texttt{304 Not Modified} when
      appropriate, reducing repeated payload bytes.

    \item \textbf{Response headers middleware.} It adds security headers
      (CSP, HSTS, etc.) and sets a short shared cache TTL suitable for profile-like
      resources.
  \end{enumerate}

  \paragraph{Semicolon-delimited \viewas{} header.}
  \viewas{} is expressed as a semicolon-delimited header that the middleware
  parses into a dict, e.g.\ \texttt{tenant\_id=acme;role=member}. This is
  intentionally simple: it is easy to generate, easy to log, and hard to
  misunderstand at 03:00.

  \begin{samepage}
  \begin{lstlisting}[style=python,caption={Control-plane middleware enforces view-as via allow-lists and resolves effective tenant scope. (Simplified from production code.)},label={lst:middleware_auth}]
# Token -> Claims -> Tenant_Id -> User_Role
tenant_id = claims["custom:tenantId"]
user_role = extract_user_role_from_groups(tenant_id, claims.get("cognito:groups"))

available = get_available_impersonations(tenant_id, user_role, claims.get("cognito:groups"))
view_as = parse_semicolon_header(event["headers"].get("X-View-As"))
if view_as:
    assert "tenant_id" in view_as and "role" in view_as
    assert view_as["tenant_id"] in [x["tenant_id"] for x in available]
    assert view_as["role"] in roles_for(view_as["tenant_id"], available)

effective_tenant = view_as["tenant_id"] if view_as else tenant_id
claims["schema_prefix"] = get_tenant_by_id(effective_tenant).schema_prefix
event["claims"] = claims
\end{lstlisting}
  \end{samepage}

  \paragraph{Policy parsing is strict.}
  Tenant policy lives in DynamoDB and includes \texttt{moduleLayout}
  fragments stored as JSON strings. During materialization, each
  fragment is parsed; invalid JSON is treated as an error (not as ``best
  effort''). This avoids serving half-valid policies that produce
  inconsistent UI composition.

% ==============================================================================
  \subsection{Edge security: middleware as first defense}
  \label{sec:edge}

  The web tier implements defense-in-depth via Next.js middleware. Before
  any server component renders, the middleware:

  \begin{enumerate}
    \item \textbf{Validates paths:} rejects path traversal attempts
      (\texttt{..}, null bytes, encoded sequences).

    \item \textbf{Applies security headers:} Content-Security-Policy,
      HSTS (production), X-Frame-Options (DENY), X-Content-Type-Options.

    \item \textbf{Checks session:} redirects unauthenticated users to
      the identity provider, preserving the original path for post-login redirect.
  \end{enumerate}

  Importantly, middleware handles \emph{authentication} and transport-level
  security; \emph{authorization} remains with the CASL-based policy model
  enforced in server layouts.

\begin{figure}[H]
  \centering
  \begin{tikzpicture}[
    layer/.style={draw, rectangle, rounded corners=3pt, minimum width=100mm,
                  minimum height=11mm, font=\small, align=center, fill=#1},
    arrow/.style={-{Latex[length=2.5mm]}, thick}
  ]
    % ==================== LAYERS (top to bottom) ====================

    % Layer 1: Edge Middleware
    \node[layer=orange!20] (l1) at (0, 4.5) {
      \textbf{Edge Middleware} --- path validation, security headers, session redirect
    };

    % Layer 2: Authentication
    \node[layer=yellow!20] (l2) at (0, 3.0) {
      \textbf{Authentication} --- JWT validation, session establishment
    };

    % Layer 3: Authorization
    \node[layer=green!20] (l3) at (0, 1.5) {
      \textbf{Authorization} --- CASL ability checks, module/route protection
    };

    % Layer 4: API Enforcement
    \node[layer=blue!15] (l4) at (0, 0) {
      \textbf{API Enforcement} --- \viewas{} validation, tenant scoping
    };

    % Layer 5: Client UX (outside trust boundary)
    \node[layer=purple!12, dashed] (l5) at (0, -1.5) {
      \textbf{Client UX} --- truthful rendering (not trusted for security)
    };

    % ==================== REQUEST FLOW ====================

    % Request arrow on left
    \draw[arrow, line width=1.2pt] (-6, 5) -- (-6, -2);
    \node[font=\scriptsize\bfseries, rotate=90] at (-6.5, 1.5) {Request Flow};

    % ==================== SECURITY BOUNDARY ====================

    % Bracket on right side
    \draw[thick, decorate, decoration={brace, amplitude=6pt, mirror}]
      (5.5, 5) -- (5.5, -0.5);
    \node[font=\scriptsize\bfseries, rotate=-90, text=gray!70!black] at (6.3, 2.25) {Security Boundary};

    % Trust boundary line
    \draw[dashed, red!50, line width=1pt] (-5.5, -0.75) -- (5.5, -0.75);
    \node[font=\tiny\bfseries, text=red!50!black, fill=white, inner sep=2pt] at (0, -0.75) {TRUST BOUNDARY};

  \end{tikzpicture}
  \caption{Defense-in-depth architecture. Each layer provides independent protection:
  Edge blocks malformed requests; Authentication validates tokens;
  Authorization enforces CASL permissions; API validates \viewas{} and scopes tenants.
  Layers 1--4 form the security boundary. The Client layer uses server-exported
  rules for truthful UX but is never trusted for authorization.}
  \label{fig:defense}
\end{figure}

% ==============================================================================
  \subsection{Web tier: Next.js as the contract consumer}
  The web tier is implemented in Next.js and serves two roles: (i) it renders
  UI, and (ii) it mediates between the browser and the control plane so
  that sensitive context (like signed \viewas{} headers) remains server-side.

  \paragraph{Session storage.}
  We use \texttt{iron-session} \cite{iron_session} to store encrypted session
  data in cookies. In practice, we keep a separation between:
  \begin{itemize}
    \item \textbf{main session:} authentication tokens and durable user
      profile,

    \item \textbf{\viewas{} session:} view-as headers (and signature), stored
      server-side and not exposed to client JavaScript.
  \end{itemize}

  \paragraph{Server-side API client.}
  The web tier uses a server API client that attaches:
  \begin{itemize}
    \item \texttt{Authorization: Bearer <JWT>}

    \item optional \texttt{X-View-As: tenant\_id=\ldots;role=\ldots}

    \item optional \texttt{X-View-As-Sig: <hmac>}
  \end{itemize}
  to calls to the control plane. The browser never needs to construct
  these headers directly.

% ==============================================================================
  \subsection{Ability endpoint: exporting truth for UX (not for security)}
  The endpoint \texttt{/api/auth/user/ability} returns packed CASL rules.
  The server constructs the ability from the authoritative session
  profile and then exports it for UI gating \cite{casl_docs}. The endpoint
  is explicitly non-cacheable to avoid stale UX when switching tenants/roles.

  \begin{samepage}
  \begin{lstlisting}[style=typescript,caption={Ability rules are built and packed on the server. The client uses them for truthful UX, not as a security gate.},label={lst:ability_endpoint}]
export const GET = withAuth(async () => {
  const packedRules = await getUserAbilityRules()
  return NextResponse.json(
    { rules: packedRules, timestamp: Date.now() },
    { headers: { 'Cache-Control': 'private, no-cache, no-store' } }
  )
})
\end{lstlisting}
  \end{samepage}

% ==============================================================================
  \subsection{Module layout endpoint: lazy config with server checks}
  The endpoint \texttt{/api/\allowbreak{}auth/\allowbreak{}user/\allowbreak{}module-layout}
  implements on-demand module config retrieval. It deep-merges tenant
  fragments over shipped defaults and applies a server-side CASL check
  before returning module-specific layouts (all modules other than \texttt{common}
  require \texttt{can(\allowbreak{}'view',\allowbreak{} module)}).

  This endpoint is where we enforce the rule:

  \begin{quote}
    \textbf{A module configuration is itself a privileged resource.}
  \end{quote}

  If the user cannot view a module, the server will not serve its
  configuration, even if the client asks nicely.

% ==============================================================================
  \subsection{Signing \viewas{}: a belt to match the suspenders}
  The web tier provides \texttt{/api/\allowbreak{}auth/\allowbreak{}view-as/\allowbreak{}sign}.
  It validates a requested tenant/role selection against \texttt{impersonatable}
  (from \texttt{/v2/me}), then generates an HMAC-SHA256 signature over
  the \texttt{X-View-As} string and stores \texttt{viewAs} and \texttt{viewAsSig}
  in the server-side session.

  This is defense-in-depth. The control plane still enforces allow-lists;
  the signature primarily protects against accidental header fabrication
  and makes tampering more detectable in logs.
