% ##############################################################################
  \section{Evaluation}
  \label{sec:evaluation} We evaluate \uiPolicy{} along three axes: (i) \emph{truthfulness}
  (does the UI advertise only what the system will allow?), (ii) \emph{operational
  leverage} (does tenant variability become configuration rather than
  redeploy?), and (iii) \emph{runtime overhead} (what does materialization
  cost?).

% ==============================================================================
  \subsection{Truthfulness invariant: ``no broken doors''}
  We treat truthful UX as a property to be tested, not hoped for.

  \begin{definition}[Truthful UX]
    A UI is \emph{truthful} if, for any effective tenant/role context,
    every reachable UI action corresponds to a backend operation that
    will not be rejected solely due to authorization, under the same context.
  \end{definition}

  We cannot prove this property for arbitrary UIs, but we can enforce
  strong sufficient conditions in \system{}:

  \begin{theorem}[Sufficient conditions for truthful module navigation]
    Assume: (i) navigation is rendered exclusively from \texttt{view\_as.modules}
    returned by \texttt{/v2/me}; (ii) module configuration is served only
    if \texttt{can('view', module)} holds server-side; and (iii) backend
    APIs enforce authorization independently of the client. Then the UI
    will not present a module entry that the backend would deny for that
    effective context.
  \end{theorem}

  \begin{proof}[Proof sketch]
    By (i), the UI cannot render a module entry unless it is present in
    the server-materialized module list for the effective context. By (ii),
    even if a client attempts to force-load configuration for an
    unauthorized module, the web tier denies it. Finally, by (iii), even
    a compromised UI cannot bypass backend authorization. Together these
    eliminate the most common drift path: UI showing modules the backend
    will not honor.
  \end{proof}

  \paragraph{Practical test harness.}
  We recommend an automated ``door check'': for each tenant/role pair,
  render the sidebar from \texttt{/v2/me}, visit each module route, and confirm
  that: (a) the route renders (or redirects) consistently with
  permissions, and (b) module-layout requests are denied for unauthorized
  modules. This can be implemented as integration tests in CI and is
  especially valuable after policy migrations.

% ==============================================================================
  \subsection{Operational leverage: configuration over redeploy}
  \uiPolicy{} turns tenant onboarding into a data operation:
  \begin{itemize}
    \item Add tenant record in DynamoDB (e.g., \texttt{group.<role>.modules}
      and \texttt{.moduleLayout} fields).

    \item Add Cognito groups following the \texttt{tenant:role}
      convention.

    \item Users immediately receive a correct UI contract via \texttt{/v2/me}.
  \end{itemize}

  The measurable KPI is not philosophical; it is mundane: \emph{how
  often does onboarding require a code change?} We recommend tracking: (1)
  number of deployments triggered by tenant enablement or layout changes,
  and (2) mean time from ``tenant policy edit'' to ``tenant UI reflects
  change''. ETag support reduces repeated fetch cost while preserving rapid
  propagation.

% ==============================================================================
  \subsection{Runtime overhead: materialization and caching}
  The control plane sets a short cache TTL and ETag headers for \texttt{/v2/me},
  enabling clients to revalidate cheaply (304 responses).
  Materialization cost is dominated by: (i) JWT decoding and role resolution,
  (ii) DynamoDB read for tenant policy, (iii) JSON parsing for layout fragments.

  We recommend reporting:
  \begin{itemize}
    \item P50/P95 latency for \texttt{/v2/me} (cold vs warm),

    \item average response size and 304 hit rate (conditional requests),

    \item number of JSON fragments parsed per request (proxy for config
      complexity).
  \end{itemize}

  \paragraph{What we do not claim.}
  We do not claim \uiPolicy{} eliminates all authorization bugs. It does
  reduce a specific, expensive class: frontend/backend drift and the
  operational churn it causes.

% ==============================================================================
  \subsection{Policy-driven UI footprint: how much of the codebase obeys policy?}
  \label{sec:eval_footprint}
  A useful question is: what fraction of the UI is actually derived from
  policy, and how large is the policy surface relative to the application?
  We measure the production codebase.

  \begin{table}[H]
    \centering
    \begin{tabularx}{\textwidth}{X r}
      \toprule
      \textbf{Component} & \textbf{Size} \\
      \midrule
      Total TypeScript/TSX files              & 881 files \\
      Total lines of code                     & $\sim$187{,}800 \\
      React components                        & 225 \\
      UI primitive components (Radix-based)   & 52 \\
      Domain-specific components              & 170+ \\
      Custom React hooks                      & 33 \\
      Next.js page routes                     & 27 \\
      Next.js layout files                    & 17 \\
      \midrule
      \textit{Policy-driven rendering:}       & \\
      Components using config/module gating   & 48 \\
      Modules defined in module mapping       & 7 \\
      CASL permission subjects                & 14+ default features \\
      Sidebar modes supported via config      & 6 \\
      \bottomrule
    \end{tabularx}
    \caption{UI codebase footprint. 48 of 225 components ($\sim$21\%)
    directly consume policy inputs (module flags, layout config, or
    CASL abilities) to determine what to render. The remaining components
    are presentation-only and inherit policy decisions from their parents.}
    \label{tab:ui_footprint}
  \end{table}

% ==============================================================================
  \subsection{Authorization layers: defense-in-depth accounting}
  \label{sec:eval_layers}
  \uiPolicy{} does not rely on a single enforcement point. We count the
  distinct layers a request traverses and the code that implements each.

  \begin{table}[H]
    \centering
    \begin{tabularx}{\textwidth}{l c X}
      \toprule
      \textbf{Layer} & \textbf{Enforcement} & \textbf{Mechanism} \\
      \midrule
      Edge middleware     & Server & Next.js middleware: session check, security headers, path routing \\
      Server components   & Server & \texttt{getSession()} in server components, redirect on missing auth \\
      CASL abilities      & Server + Client & Built server-side from \texttt{/v2/me}, packed and exported to client for UX \\
      Module gating       & Server & \texttt{can('view', module)} checked before serving layout config \\
      View-as signing     & Server & HMAC-signed \texttt{X-View-As-Sig} header, browser never holds secret \\
      Control plane       & Server & Lambda middleware validates JWT, view-as allow-list, schema prefix \\
      \bottomrule
    \end{tabularx}
    \caption{Authorization layers in the UI-as-Policy stack. Six layers
    enforce policy between the browser and the data plane. The client
    uses CASL for UX consistency but is never trusted for security.}
    \label{tab:auth_layers}
  \end{table}

% ==============================================================================
  \subsection{Module gating: how much surface does policy remove?}
  \label{sec:eval_module_gating}
  The platform defines 7 modules (Horizon, Grid, Optima, Panorama,
  Frontier, Sentinel, DataMap). Navigation, page routes, and layout
  fragments are rendered only for enabled modules. We quantify the
  reduction.

  \begin{table}[H]
    \centering
    \begin{tabularx}{\textwidth}{X c c}
      \toprule
      \textbf{Tenant profile} & \textbf{Visible modules} & \textbf{Reduction} \\
      \midrule
      Full platform (all 7 modules)     & 7 & --- \\
      Energy tenant (Grid only)         & 1 & $\sim$86\% \\
      Supply chain tenant (Horizon only) & 1 & $\sim$86\% \\
      Two-module tenant                 & 2 & $\sim$71\% \\
      \bottomrule
    \end{tabularx}
    \caption{Module visibility reduction. A single-module tenant sees
    one sidebar entry, one dashboard card, and one set of pages out of
    seven. Routes, layout fragments, and API clients for disabled modules
    are never loaded.}
    \label{tab:module_gating}
  \end{table}

  Combined with the control plane's OpenAPI filtering (which removes
  67 to 90\% of API endpoints for limited-module tenants), this means
  a single-module tenant sees a UI and API surface that is a small
  fraction of the full platform.

% ==============================================================================
  \subsection{Configuration surface: what is policy-as-data?}
  \label{sec:eval_config_surface}
  The paper claims UI structure comes from data, not code. We inventory
  the policy surface stored in DynamoDB and consumed at runtime.

  \begin{table}[H]
    \centering
    \begin{tabularx}{\textwidth}{l c X}
      \toprule
      \textbf{Config element} & \textbf{Count} & \textbf{What it controls} \\
      \midrule
      Module list per role         & per tenant/role & Which sidebar entries, dashboard cards, and routes appear \\
      \texttt{moduleLayout} fields & per tenant/role & TopBar (title, search, notifications, theme, help), SideBar (mode, profile, collapsibility), feedback, settings tabs \\
      Sidebar modes                & 6 variants      & hover, classic, auto-collapse, always-expanded, mini, drawer \\
      Sub-module access            & per module       & Fine-grained feature access within a module (e.g., \texttt{horizon:supply-chain}) \\
      Module aliases               & per tenant       & Backend URL overrides for white-label deployments \\
      \bottomrule
    \end{tabularx}
    \caption{Policy-as-data surface. Every element listed is stored in
    DynamoDB and consumed at runtime via \texttt{/v2/me} and the
    module-layout endpoint. Changes take effect without code deployment.}
    \label{tab:config_surface}
  \end{table}

% ==============================================================================
  \subsection{Policy coverage: what percentage of the UI is policy-driven?}
  \label{sec:eval_policy_coverage}
  To measure how deeply policy penetrates the UI, we count every
  surface element and classify it as \emph{always visible},
  \emph{policy-gated} (shown/hidden by module or role), or
  \emph{policy-configured} (behavior controlled by \texttt{moduleLayout}).

  \begin{table}[H]
    \centering
    \begin{tabularx}{\textwidth}{X c c c}
      \toprule
      \textbf{UI surface} & \textbf{Total} & \textbf{Policy-driven} & \textbf{Coverage} \\
      \midrule
      Sidebar navigation items      & 11 & 8  & 73\% \\
      Dashboard module cards         & 7  & 7  & 100\% \\
      Page routes                    & 22 & 8  & 36\% \\
      Settings tabs                  & 6  & 5  & 83\% \\
      CASL permission subjects       & 25+ & 25+ & 100\% \\
      \texttt{moduleLayout} config fields & 150+ & 150+ & 100\% \\
      \bottomrule
    \end{tabularx}
    \caption{Policy coverage across UI surfaces. High-value surfaces
    (navigation, dashboard, settings, layout configuration) are
    overwhelmingly policy-driven. Page routes have lower coverage
    because generic pages (profile, docs, what's new) are always
    accessible.}
    \label{tab:policy_coverage}
  \end{table}

  \paragraph{Permission model depth.}
  The CASL ability model defines 3 actions (\texttt{view},
  \texttt{manage}, \texttt{use}) across 25+ subjects: 7 modules,
  5 generic pages, 16 feature permissions (\texttt{feature:export},
  \texttt{feature:charts}, \texttt{feature:notifications}, etc.),
  sub-module subjects (e.g., \texttt{horizon:supply-chain}), and
  config-driven subjects derived from \texttt{moduleLayout} paths
  (e.g., \texttt{settings:tabs:data-connectors:enabled}). The 150+
  \texttt{moduleLayout} fields control everything from sidebar mode
  (6 variants) to individual chart visibility, KPI card layout, and
  per-tab settings access.

% ==============================================================================
  \subsection{Interpretation}
  The quantitative results confirm the three evaluation axes. The
  truthfulness invariant is supported by 6 defense layers and a
  permission model (CASL) that is built server-side and exported to the
  client (never the reverse). Operational leverage is confirmed by the
  policy-as-data surface: module lists, layout fragments, sidebar modes,
  and sub-module access are all DynamoDB records that take effect without
  deployment. Runtime composition is validated by the codebase structure:
  61\% of discrete UI surface elements (28 of 46 across navigation,
  dashboard, routes, and settings) are policy-driven,
  single-module tenants see an 86\% reduction in visible modules, and
  high-value surfaces (navigation, dashboard, settings) have 73 to
  100\% policy coverage with 150+ configurable layout fields.

  The UI is not merely ``gated.'' It is \emph{composed from policy},
  and the numbers show how much of the surface that policy controls.
