% ##############################################################################
  \section{Related Work}
  \label{sec:relatedwork}
  \paragraph{ABAC, RBAC, and policy expressiveness.}
  ABAC formalizes authorization decisions based on attributes rather than
  static roles and is a standard reference point \cite{nist_abac}.
  \uiPolicy{} is compatible with ABAC: the server may compute abilities
  using attributes, but the key contribution here is the UI
  contract that prevents the frontend from guessing.

  \paragraph{Feature toggles and product variability.}
  Feature flags separate release from deployment and are widely used for
  rollout control \cite{fowler_feature_toggles}. However, feature flags alone
  do not guarantee UI truthfulness: they can become a parallel universe
  of conditions that diverge from authorization.\@ \uiPolicy{} can be
  viewed as feature toggles with teeth: configuration is scoped by tenant/role
  and validated against the same ability model.

  \paragraph{Policy engines and external authorization services.}
  OPA provides a general-purpose policy engine and a declarative
  language to offload authorization decisions \cite{opa_docs}. Cedar (used
  by Amazon Verified Permissions) provides a language for authorization policies
  and standardized evaluation models \cite{cedar_docs,aws_verified_permissions}.
  These systems are highly relevant when the goal is centralized
  authorization logic.\@ \uiPolicy{} is orthogonal: we focus on how policy
  decisions become UI structure in a multi-tenant product.

  \paragraph{Relationship-based authorization and global systems.}
  Zanzibar describes a global, consistent authorization system used
  across Google services \cite{zanzibar}. While \system{} does not
  implement a Zanzibar-style relationship store, Zanzibar underscores an
  important theme: authorization becomes a distributed-systems problem
  at scale.\@ \uiPolicy{} addresses a nearby problem: when authorization scales,
  the UI must not become an unreliable narrator.

  \paragraph{Impersonation and operational risk.}
  OWASP explicitly highlights account impersonation as a security risk in
  citizen development contexts \cite{owasp_impersonation}. Our \viewas{}
  design aligns with the recommended mitigations: explicit actor
  attribution, guardrails, and defense-in-depth.
