% ##############################################################################
  \section{Limitations and Future Work}
  \label{sec:futurework} \uiPolicy{} makes a specific trade: it shifts some product definition into
  data. This buys flexibility, but it introduces new responsibilities.

  \paragraph{Policy schema validation.}
  Today, malformed \texttt{moduleLayout} JSON is rejected at runtime.
  Future work should include:
  \begin{itemize}
    \item a versioned JSON schema for layout fragments,

    \item offline validation tooling (CI + pre-commit),

    \item migration helpers when layout structures evolve.
  \end{itemize}

  \paragraph{Signature verification in the control plane.}
  The web tier signs \texttt{X-View-As}. Verifying \texttt{X-View-As-Sig}
  in the control plane would strengthen defense-in-depth and make header
  provenance auditable end-to-end.

  \paragraph{Push-based invalidation.}
  ETags and short TTL caching work well for most edits, but critical
  policy updates may benefit from push invalidation (e.g., event-driven cache
  busting or policy version bumps in the session).

  \paragraph{Finer-grained capabilities.}
  Today, capabilities are primarily module-centric. Extending to feature-
  and resource-level controls (within a module) is feasible, but must preserve
  the single-source-of-truth property and avoid reintroducing drift through
  ad-hoc UI conditions.

  \paragraph{Governance and authoring experience.}
  A policy-as-data system needs governance: who can change tenant policy,
  how changes are reviewed, and how rollbacks work. A structured policy
  editor (with preview and diffing) is a natural next step.
