\section{Architecture Overview}
\label{sec:architecture}

Figure~\ref{fig:arch_overview} shows the boundary. The control plane combines API Gateway, Lambda, DynamoDB, Cognito validation, and a small middleware spine. It owns the decision path. The data plane combines EKS services, module Lambdas, Fargate jobs, Airflow-driven workflows, and tenant-scoped storage. It owns execution.

\begin{figure}[H]
  \centering
  \resizebox{0.86\textwidth}{!}{%
  \begin{tikzpicture}[
    node distance=13mm and 20mm,
    box/.style={draw, rounded corners=3pt, align=center, minimum width=33mm, minimum height=14mm, font=\small\sffamily},
    arrow/.style={-{Latex[length=2.4mm]}, thick, color=gray!75},
    label/.style={font=\scriptsize\sffamily, fill=white, inner sep=2pt, align=center}
  ]
    \node[box, fill=yellow!18, draw=yellow!55!black] (idp)
      {\textbf{Identity Provider}\\Cognito\\JWT + groups};
    \node[box, fill=cyan!12, draw=cyan!60!black, right=22mm of idp] (web)
      {\textbf{Web Tier}\\Next.js\\server session};
    \node[box, fill=blue!12, draw=blue!55!black, below=16mm of web] (cp)
      {\textbf{Control Plane}\\API Gateway + Lambda\\middleware spine};
    \node[box, fill=violet!12, draw=violet!55!black, left=22mm of cp] (policy)
      {\textbf{Policy Store}\\DynamoDB\\tenants + modules};
    \node[box, fill=green!12, draw=green!55!black, right=22mm of cp] (dp)
      {\textbf{Data Plane}\\EKS + Fargate\\services + jobs};
    \node[box, fill=blue!8, draw=blue!55!black, below=16mm of cp] (lambda)
      {\textbf{Module Lambdas}\\Analytics / Query\\scoped requests};
    \node[box, fill=violet!12, draw=violet!55!black, below=16mm of lambda] (db)
      {\textbf{Tenant DB}\\RDS Proxy $\rightarrow$ Postgres\\\texttt{\{prefix\}\_core}};

    \draw[arrow] (web.west) -- node[label,above] {authenticates} (idp.east);
    \draw[arrow] (web.south) -- node[label,right] {\code{/v2/me} + ETag} (cp.north);
    \draw[arrow] (cp.west) -- node[label,above] {tenant lookup} (policy.east);
    \draw[arrow] (cp.east) -- node[label,above] {scoped calls} (dp.west);
    \draw[arrow] (cp.south) -- node[label,right] {injects scope} (lambda.north);
    \draw[arrow] (lambda.south) -- node[label,right] {sets \code{search\_path}} (db.north);
    \draw[arrow] (dp.south east) -- node[label,right] {scoped compute} (db.north east);

    \node[below=7mm of db, text width=130mm, align=center, font=\small\sffamily]
      {\textbf{Boundary rule.} Tenancy and capability decisions originate in the control plane. Execution systems receive explicit scope and run workloads. They do not rediscover policy.};
  \end{tikzpicture}%
  }
  \caption{The control-plane/data-plane boundary. The control plane resolves identity, effective tenant context, modules, impersonation, API surface, and downstream scope. The data plane executes within that injected scope.}
  \label{fig:arch_overview}
\end{figure}

\subsection{The \code{/v2/me} contract}

The control plane exposes a versioned identity contract that the web tier calls before page rendering. It is similar in spirit to an OpenID Connect UserInfo endpoint \cite{oidc_core}, but it carries platform-specific tenant and module state.

A simplified response is shown below.

\begin{lstlisting}[style=json, caption={Simplified identity and effective-context contract.}]
{
  "actor": {
    "email": "ada@acme.example",
    "tenant_id": "platform-a",
    "user_role": "admin"
  },
  "view_as": {
    "tenant_id": "acme",
    "user_role": "member",
    "schema_prefix": "acme"
  },
  "impersonatable": [
    { "tenant_id": "acme", "roles": ["demo", "member", "admin"] }
  ]
}
\end{lstlisting}

The important field is \scopeprefix{}. It appears under \viewas{} because database scope follows the effective tenant, not the support engineer's home tenant. When no impersonation is active, \viewas{} simply mirrors the actor's own tenant context.

The endpoint uses ETag and \code{If-None-Match} semantics from RFC 9110 \cite{rfc9110}. On a cache hit, the Lambda returns \code{304 Not Modified} with no body. The fingerprint covers identity and role fields; module changes are operator actions and take effect at the next full materialization. That trade-off gives the platform cheap, frequent revalidation without pretending policy changes are end-user events.

\subsection{What belongs on each side}

The boundary is practical, not theological. Authentication, tenant lookup, module gating, impersonation validation, API-surface filtering, and scope derivation live in the control plane. Query execution, model execution, workflow orchestration, data transformation, exports, and customer-facing computations live in the data plane.

The architecture becomes easier to reason about once that split is enforced. When a data-plane service receives a request, it should not ask, ``which tenant am I really serving?'' It should already know. The control plane has answered that question upstream.
