% ##############################################################################
  \section{Control Plane Design}
  \label{sec:control_plane_design} The control plane is the platform's \emph{governance
  cortex}: it authenticates, derives effective tenant context, materializes
  module availability, and produces contracts that downstream components
  can execute without guessing. The guiding rule is simple:

  \begin{quote}
    \textbf{Downstream components may execute. They may not improvise tenancy.}
  \end{quote}

% ==============================================================================
  \subsection{Endpoint taxonomy: contracts, routers, and delegation}
  The control plane exposes three categories of endpoints:

  \begin{itemize}
    \item \textbf{Context contracts (identity \& scoping):} \texttt{/v2/me}
      returns immutable actor identity, effective context (including \viewas{}),
      and derived scoping signals such as \texttt{schema\_prefix}.

    \item \textbf{Surface description:} \texttt{/v1/openapi.json} returns
      an OpenAPI document filtered so clients only see endpoints corresponding
      to enabled modules.

    \item \textbf{Module routers:} \texttt{/v1/horizon/*}, \texttt{/v1/grid/*},
      and \texttt{/v1/datamap/*} implement module APIs guarded by module
      availability checks.
  \end{itemize}

  Critically, some module routers in the control plane act as \emph{delegators}:
  they forward requests to module-specific Lambda backends (e.g., Horizon
  and Grid) while injecting derived scoping values. This makes the control
  plane the only place where tenant scope is \emph{decided}, even when the
  work is done elsewhere.

% ==============================================================================
  \subsection{Middleware spine: three decorators that do 80\% of the job}
  The control plane uses Powertools' REST resolver with a small but
  decisive middleware chain \cite{aws_powertools}. The order matters:

  \begin{samepage}
    \begin{lstlisting}[style=python,caption={Middleware chain used by the control plane Lambda. Auth runs first; response shaping and conditional requests run around the handler.},label={lst:cp_decorators}]
@auth_middleware
@etag_middleware
@response_headers_middleware
@logger.inject_lambda_context(...)
def lambda_handler(event, context):
    return app.resolve(event, context)
\end{lstlisting}
  \end{samepage}

  \paragraph{Auth middleware (truth injection).}
  \texttt{auth\_middleware} validates JWTs, derives \texttt{tenant\_id}
  and role from Cognito groups, computes an explicit allow-list of impersonation
  targets, and parses/validates the semicolon-delimited \viewas{} header:

  \begin{center}
    \texttt{X-View-As: tenant\_id=<id>;role=<role>[;mode=<...>]}
  \end{center}

  If \viewas{} is present, it is validated against \texttt{available\_impersonations}
  before it is accepted. The middleware then resolves the tenant record
  and injects \texttt{schema\_prefix} into the request claims.

  This is a deliberate design choice: downstream components should never
  accept \texttt{schema\_prefix} (or equivalent scoping) from the client.
  That is the confused-deputy trap.

  \paragraph{Identity derivation: the \texttt{tenant\_id:role}
  convention.}
  Cognito user-pool groups follow a strict naming convention: \texttt{tenant\_id:role}
  (e.g., \texttt{causify:admin}, \texttt{acme:member}). The auth middleware
  splits each group on \texttt{:}, filters for the requested tenant,
  extracts the role portion, and selects the highest-precedence role when
  a user belongs to multiple groups for the same tenant.

  \begin{samepage}
    \begin{lstlisting}[style=python,caption={Role derivation from Cognito groups: groups follow a \texttt{tenant\_id:role} convention and are resolved by precedence.},label={lst:group_parsing}]
valid_groups = [g for g in user_groups if ":" in g
                and g.startswith(f"{tenant_id}:")]
user_roles = [g.split(":")[1] for g in valid_groups]
order = ["admin", "member", "demo"]
user_roles.sort(key=lambda r: order.index(r) if r in order else len(order))
user_role = user_roles[0]
\end{lstlisting}
  \end{samepage}

  This convention is load-bearing: it lets the control plane derive both
  tenant identity and role from a single JWT claim without an external policy
  lookup, and it makes the impersonation allow-list computable from the
  same group list.

  \paragraph{Response headers middleware (hardening + ETag generation).}
  \texttt{response\_headers\_middleware} adds baseline security headers
  (\texttt{X-Frame-Options: DENY}, strict CSP, HSTS, etc.) and sets a short
  caching directive (\texttt{Cache-Control: public, max-age=300}). It also
  computes an \texttt{ETag} as an MD5 hash over the response body. (The
  ETag is quoted per HTTP conventions.)

  \paragraph{ETag middleware (cheap revalidation).}
  \texttt{etag\_middleware} implements conditional requests: when
  \texttt{If-None-Match} matches the current response body hash, the
  handler returns \texttt{304 Not Modified} with the same standard headers.
  This makes the platform's most frequently asked question (``who am I,
  and what tenant am I in?'') cheap \cite{rfc9110}.

  If the UI asks this endpoint a lot, that's not a bug. That's anxiety. We
  treat it like a first-class workload.

  \paragraph{Request lifecycle.}
  Figure~\ref{fig:request_lifecycle} makes the middleware execution
  order concrete: a request passes through authentication, conditional
  caching, and response hardening before reaching module routing.

  \begin{figure}[H]
    \centering
    \begin{cleanpic}
      [ mwstep/.style={cbox, align=left, text width=0.82\textwidth, minimum height=10mm},
      mwinner/.style={cbox, align=left, text width=0.74\textwidth, minimum height=10mm},
      mwcontainer/.style={draw=cpblue!50, rounded corners=5pt, inner sep=10pt, fill=cpblue!6, minimum width=\dimexpr0.82\textwidth+14pt\relax, drop shadow={shadow xshift=0.6mm, shadow yshift=-0.6mm, opacity=0.3}},
      ] \node[mwstep, fill=webteal!25, draw=webteal!50] (s1)
      {\textbf{1. Client} sends request to API Gateway};

      % --- Middleware group (three sections inside one container) ---
      \node[mwinner, fill=idpyellow!28, draw=idpyellow!55, below=14mm of
      s1] (mw1)
      {\textbf{auth\_middleware}: validate JWT, derive tenant/role, parse \viewas{}, validate impersonation, inject \texttt{schema\_prefix}};
      \node[mwinner, fill=cpblue!25, draw=cpblue!50, below=3mm of mw1] (mw2)
      {\textbf{etag\_middleware}: check \texttt{If-None-Match}; if match $\rightarrow$ return \texttt{304}};
      \node[mwinner, fill=cpblue!25, draw=cpblue!50, below=3mm of mw2] (mw3)
      {\textbf{response\_headers\_middleware}: add security headers (\texttt{X-Frame-Options}, CSP, HSTS), compute \texttt{ETag}};
      \node[font=\scriptsize\sffamily\bfseries, text=cpblue!80!black, anchor=south
      west] at ([shift={(2pt,5pt)}]mw1.north west) (mwtitle) {2. Middleware};

      % --- Middleware container on background ---
      \begin{pgfonlayer}{background}
        \node[mwcontainer, fit=(mwtitle)(mw1)(mw2)(mw3)] (mwbox) {};
      \end{pgfonlayer}

      % --- Remaining steps ---
      \node[mwstep, fill=cpblue!25, draw=cpblue!50, below=14mm of mw3] (s3)
      {\textbf{3. Router}: match path to handler}; \node[mwstep, fill=dpgreen!25,
      draw=dpgreen!50, below=4mm of s3] (s4) {\textbf{4. Module Guard}: check module availability for effective context};
      \node[mwstep, fill=dpgreen!25, draw=dpgreen!50, below=4mm of s4] (s5)
      {\textbf{5. Delegate} to Module Lambda with injected scope};

      % --- Arrows ---
      \draw[carrow] (s1) -- (mwbox.north); \draw[carrow] (mw1) -- (mw2);
      \draw[carrow] (mw2) -- (mw3); \draw[carrow] (mwbox.south) -- (s3);
      \draw[carrow] (s3) -- (s4); \draw[carrow] (s4) -- (s5);
    \end{cleanpic}
    \caption{Request lifecycle through the control plane middleware
    spine. Auth runs first (truth injection), followed by conditional
    caching, response hardening, routing, module gating, and delegation.}
    \label{fig:request_lifecycle}
  \end{figure}

% ==============================================================================
  \subsection{Module availability as a first-class primitive}
  Module availability is computed from tenant configuration stored in
  DynamoDB \cite{dynamodb_docs}. The function \texttt{get\_available\_modules(tenant\_id,
  user\_role)} returns:

  \begin{itemize}
    \item \texttt{modules}: enabled module identifiers (e.g., \texttt{horizon},
      \texttt{grid}, \texttt{datamap})

    \item \texttt{submodules}: enabled sub-module identifiers and their parent
      mapping (e.g., \texttt{horizon: supply-chain})

    \item \texttt{module\_layout}: module-specific JSON fragments stored
      as JSON strings in DynamoDB and parsed server-side (strict parsing;
      invalid JSON fails fast).
  \end{itemize}

  This strict parsing matters operationally: configuration that silently
  half-loads is worse than configuration that fails loudly. Loud failures
  are debuggable.

% ==============================================================================
  \subsection{Surface minimization: filtered OpenAPI, not just forbidden
  endpoints}
  Most APIs only deny. The control plane also \emph{stops advertising}.
  The OpenAPI endpoint generates the full schema and then filters it:

  \begin{itemize}
    \item Always allow \texttt{user} and \texttt{tenant} tags.

    \item Allow module endpoints only when the endpoint tags intersect enabled
      modules.

    \item Remove unused component schemas after filtering.
  \end{itemize}

  \begin{samepage}
    \begin{lstlisting}[style=python,caption={OpenAPI filtering logic: paths are kept if their tags are default-allowed or enabled by module availability.},label={lst:openapi_filter}]
default_allowed_tags = ["user", "tenant"]
is_default_allowed = any(tag in default_allowed_tags for tag in operation_tags)
is_module_allowed  = any(tag in available_modules   for tag in operation_tags)
if is_default_allowed or is_module_allowed:
    keep(path, method)
\end{lstlisting}
  \end{samepage}

  This reduces surface drift: the platform's published interface becomes
  a function of tenant capability, not a static document with wishful
  thinking.

% ==============================================================================
  \subsection{Module guards: authorization at router boundaries}
  Module routers are protected by a \texttt{@requires\_module(module)} decorator.
  The decorator evaluates module availability for the \emph{effective
  context} (actor or \viewas{}) and raises a forbidden error when a
  module is not enabled.

  This is intentionally blunt: it is not trying to infer fine-grained permissions.
  It ensures the coarse-grained capability boundary is enforced
  everywhere, including at delegating routers.

  While \texttt{@requires\_module} checks the parent module, a tenant may
  have that module enabled while only some \emph{submodules} are enabled:
  core product surfaces, optional add-ons, or sub-features rolled out in
  stages. For a finer gate, routes use
  \texttt{@requires\_submodule("module/submodule")} (for example
  \texttt{@requires\_submodule("horizon/cpg")} for CPG submodule APIs under
  Horizon). The decorator evaluates the effective context against
  \texttt{get\_available\_modules(tenant\_id, user\_role)}, which returns
  tenant \texttt{submodules} from DynamoDB together with \texttt{modules};
  the submodule must be allowed even when the parent module is already on.

  \begin{samepage}
    \begin{lstlisting}[style=python,caption={Module guard enforces capability boundaries using the effective tenant/role context. Submodule routes add \texttt{@requires\_submodule}.},label={lst:requires_module}]
@requires_module("horizon")
def list_kpis(...):
    return invoke_horizon_lambda("/kpis", query_params)

@requires_module("horizon")
@requires_submodule("horizon/cpg")
def list_cpg_connectors(...):
    return invoke_horizon_lambda("/cpg/connectors", query_params)

\end{lstlisting}
  \end{samepage}

% ==============================================================================
  \subsection{Delegation with scope injection: control plane as a tenancy
  firewall}
  When module routers delegate to module-specific backends (Horizon/Grid
  Lambdas), the control plane injects \texttt{schema\_prefix} into query
  parameters:

  \begin{samepage}
    \begin{lstlisting}[style=python,caption={Delegation pattern: the control plane injects derived tenant scope into downstream module backends.},label={lst:schema_prefix_injection}]
claims = get_current_event().get("claims")
schema_prefix = claims.get("schema_prefix")
query_params["schema_prefix"] = schema_prefix
response = invoke_lambda(horizon_lambda_function, path, query_params, method, body)
\end{lstlisting}
  \end{samepage}

  This protects against the most common tenant-isolation footgun: a
  downstream service that trusts client-provided scoping parameters. Here,
  the client never gets the chance.

  \paragraph{Mechanics: constructing an API Gateway event.}
  The delegation is not an HTTP call. The control plane constructs an
  API Gateway--compatible event and invokes the downstream Lambda synchronously
  via \texttt{boto3}:

  \begin{samepage}
    \begin{lstlisting}[style=python,caption={Lambda-to-Lambda delegation: the control plane builds an API Gateway--like event with injected scope and invokes the module backend synchronously.},label={lst:lambda_invoke}]
lambda_event = {
    "httpMethod": method,
    "path": path,
    "queryStringParameters": query_params,  # schema_prefix is here
    "headers": {"Content-Type": "application/json"},
    "body": json.dumps(body) if body else None,
}
response = boto3.client("lambda").invoke(
    FunctionName=function_name,
    InvocationType="RequestResponse",
    Payload=json.dumps(lambda_event),
)
\end{lstlisting}
  \end{samepage}

  This design means that module backends never need their own API Gateway
  deployment: they receive the same event shape they would get from a
  public gateway, but the caller is always the control plane, and the
  scope is always server-derived. It also keeps the delegation path testable:
  any event that matches the shape can be used in integration tests
  without an actual gateway.
