\section{Security Model}
\label{sec:security}

The architecture defends against four attack and failure classes: forged tenancy context, confused-deputy scope substitution, surface drift that exposes disabled endpoints, and unattributed operator action during impersonation.

\subsection{Layered controls}

The deployed system uses five layers.

\begin{enumerate}
  \item \textbf{JWT validation.} A valid Cognito JWT is required, but token claims alone do not decide tenant scope. The middleware re-derives membership from the platform table.
  \item \textbf{View-as guardrails.} Impersonation requires a server-side allow-list match, a signed payload, a live session, and a registered nonce.
  \item \textbf{Scope injection.} Downstream services receive \scopeprefix{} from the control plane and reject calls that omit it.
  \item \textbf{Surface minimization.} Unlicensed routes are missing from the served API spec and return \code{403} if called directly.
  \item \textbf{Network and transport hardening.} Data stores and compute services are not exposed as public shortcuts around the control plane.
\end{enumerate}

No one layer is treated as magic. A valid JWT does not grant arbitrary scope. A visible UI control does not authorize a backend call. A syntactically valid schema prefix does not prove tenant membership. Each layer narrows a different failure mode.

\subsection{What remains risky}

The policy store becomes a critical surface. Moving policy from code to data is useful, but it changes the blast radius of a bad write. A faulty tenant record can break module access and scope for that tenant. A malicious write could do worse.

The mitigation is operational discipline: scoped IAM writes, reviewable changes, strict parsing, schema validation, audit logs, and a habit of treating policy updates with the same seriousness as code deploys. Configuration is not casual just because it is stored in DynamoDB.

The second residual risk is wrong derivation inside the control plane. If the control plane derives the wrong tenant scope and all logged fields agree with that wrong value, downstream services cannot detect the mistake. Section~\ref{sec:evaluation} discusses the detection gap honestly.
