\section{Operations and Reliability}
\label{sec:operations}

The control-plane boundary helped most in the ordinary work of operating tenants. It made changes smaller, logs clearer, and failure modes less mysterious.

\subsection{Lifecycle operations}

Most tenant lifecycle changes are data changes:

\begin{itemize}
  \item add a tenant,
  \item enable or disable a module,
  \item change module layout,
  \item grant or revoke an impersonation target,
  \item change a tenant schema prefix.
\end{itemize}

These updates touch the tenant policy store and are visible at the next full materialization. Adding a new module type still requires code: routes, handlers, decorators, defaults, tests, and documentation. That is the right split. Routine tenant operations should not redeploy 12 services. New product surface should still go through engineering review.

\subsection{ETag revalidation}

The \code{/v2/me} contract is called often enough that performance matters. Revalidating identity and effective context on every page load is valuable; re-sending the full payload each time is wasteful. ETag-based conditional requests keep the contract cheap enough to use as the platform's source of truth.

The steady-state hit rate is 91\%. A \code{304 Not Modified} response returns at 68 ms P50. Full materialization is slower, but less frequent. That is a reasonable trade: correctness is checked often, while expensive identity assembly happens when it has something to say.

\subsection{Logs and incident handling}

Every request logs actor tenant, effective tenant, module context, and \scopeprefix{} where applicable. The distinction between actor and effective tenant is not a compliance flourish; it is how support sessions remain explainable.

A representative incident query looks for mismatches between actor tenant, effective tenant, and impersonation state. That catches missing-prefix requests and non-impersonation context divergence. It does not catch a wrong value that was consistently derived upstream. The logs are useful, not omniscient.

\subsection{Adoption path}

Migration does not need a flag day. The practical sequence is:

\begin{enumerate}
  \item deploy the control plane Lambda, tenant table, and \code{/v2/me} endpoint,
  \item route selected service calls through the control plane,
  \item replace local middleware with shared context handling,
  \item add module decorators to routes,
  \item move per-service module flags into the central tenant record,
  \item turn on OpenAPI filtering for tenants with limited modules.
\end{enumerate}

The migration of 12 services took approximately 160 engineer-hours across two sprints, including tests. That is not free. It is still cheaper than carrying a permanent governance tax in every service.
