% ##############################################################################
  \section{Evaluation}
  \label{sec:evaluation} This section evaluates whether the control-plane/data-plane
  boundary achieves its goals: explicit tenancy, minimized surface drift,
  operational clarity, and acceptable overhead. We report evaluation as
  a set of measurable questions with concrete metrics where available.

% ==============================================================================
  \subsection{Q1: Is the control plane fast enough for interactive use?}
  The web tier calls \texttt{/v2/me} frequently. The control plane should
  keep that contract cheap; here we report \texttt{/v2/me} latency from a
  load of 100 total requests with concurrency~10, as summary statistics
  (Table~\ref{tab:cp_perf_summary}) and percentiles
  (Table~\ref{tab:cp_perf_percentiles}).

  \begin{table}[H]
    \centering
    \begin{tabular}{lr}
      \toprule \textbf{Metric} & \textbf{Value (ms)} \\
      \midrule
      Mean    & 1415.15 \\
      Std Dev & 372.52 \\
      Min     & 1001.21 \\
      Max     & 2594.79 \\
    \end{tabular}
    \caption{\texttt{/v2/me} latency: summary statistics (100 requests, concurrency~10).}
    \label{tab:cp_perf_summary}
  \end{table}

  \begin{table}[H]
    \centering
    \begin{tabular}{lr}
      \toprule \textbf{Percentile} & \textbf{Value (ms)} \\
      \midrule
      p1  & 1008.74 \\
      p5  & 1025.96 \\
      p10 & 1042.47 \\
      p25 & 1077.33 \\
      p50 & 1498.69 \\
      p75 & 1541.80 \\
      p90 & 1842.08 \\
      p95 & 2341.29 \\
      p99 & 2510.98 \\
      \bottomrule
    \end{tabular}
    \caption{\texttt{/v2/me} latency: percentiles (same run as Table~\ref{tab:cp_perf_summary}).}
    \label{tab:cp_perf_percentiles}
  \end{table}

% ==============================================================================
  \subsection{Q2: Does scope injection prevent tenant confusion?}
  We treat ``tenant confusion'' as an incident class: any request where a
  component touches data outside the effective tenant scope.

  We validate scope injection in two ways:
  \begin{itemize}
    \item \textbf{Static checks:} module routers require \texttt{schema\_prefix}
      and refuse to operate without it.

    \item \textbf{Runtime correlation:} logs include actor tenant, effective
      tenant, and \texttt{schema\_prefix}; anomalies are detectable via
      simple queries (e.g., schema\_prefix not matching effective tenant).
  \end{itemize}

  \paragraph{Metric.}
  Count occurrences of mismatched scope fields across the request path. The
  target is zero; the practical goal is ``approaches zero and spikes are
  explainable.''

  \paragraph{Code evidence.}
  The injection chain is traceable in the codebase.
  \texttt{auth\_middleware} (middleware.py) derives \texttt{schema\_prefix}
  from the tenant record, not from any client-supplied header or query
  parameter. The controller injects this value into every downstream Lambda
  invocation: \texttt{services/horizon.py} and \texttt{services/grid.py}
  both pass it as a required parameter in the internal request payload.
  Horizon Lambda refuses to operate without \texttt{schema\_prefix},
  raising a \texttt{BadRequestError} if the value is missing. This means
  a confused-deputy scenario requires \emph{two} independent failures:
  the controller must mis-derive the prefix \emph{and} the module Lambda
  must accept it.

% ==============================================================================
  \subsection{Q3: Does OpenAPI filtering reduce surface drift?}
  Surface drift occurs when clients can discover or call endpoints for
  modules that are disabled. We measure:
  \begin{itemize}
    \item number of module endpoints present in the raw OpenAPI vs filtered
      OpenAPI,

    \item attempted calls to disabled module endpoints (should be denied
      and ideally rare),

    \item UI-to-spec alignment: endpoints used by the client must exist
      in the filtered spec.
  \end{itemize}

  \paragraph{Result shape.}
  In a healthy system, filtered OpenAPI removes a large fraction of endpoints
  for tenants with limited modules, and attempted calls to disabled
  endpoints remain near zero (ideally because clients never see them).

  \paragraph{Code evidence.}
  Two functions in \texttt{services/openapi.py} implement the filtering.
  \texttt{filter\_openapi\_paths()} removes every path whose tags do not
  intersect the tenant's enabled module set.
  \texttt{filter\_openapi\_schemas()} then recursively prunes schema
  definitions that are no longer referenced by any surviving path, preventing
  information disclosure about features the tenant has not licensed.
  Critically, the filtering uses the \emph{effective} context (the
  \texttt{view\_as} identity after impersonation resolution), not a static
  configuration file. This means the same deployment serves different API
  surfaces to different tenants without a redeploy, and an impersonating
  support agent sees exactly the surface available to the tenant they are
  assisting.

% ==============================================================================
  \subsection{Q4: Is the control plane actually ``thin''?}
  The architecture claims that governance is concentrated into a thin
  layer. We validate this by measuring the control plane's code footprint
  and comparing it to the data plane infrastructure it governs.

  \begin{table}[H]
    \centering
    \begin{tabularx}{\textwidth}{X r}
      \toprule
      \textbf{Component} & \textbf{Size} \\
      \midrule
      Total control plane Python files     & 37 files \\
      Total control plane lines of code    & $\sim$6{,}174 \\
      Middleware stack (\texttt{core/middleware.py}) & 248 lines \\
      Auth modules (JWT + decorators)      & 133 lines \\
      All six routers combined             & 2{,}420 lines \\
      Pydantic data models                 & 75 model classes \\
      \midrule
      \textit{Data plane for comparison:}  & \\
      Kubernetes YAML manifests            & 743 files \\
      Kubernetes namespaces / applications & 12 \\
      Terraform infrastructure modules    & 18 \\
      \bottomrule
    \end{tabularx}
    \caption{Control plane code footprint vs.\ data plane infrastructure.
    The entire governance layer (authentication, impersonation, scope
    injection, module gating, surface filtering, and response
    hardening) fits in $\sim$6{,}200 lines of Python, governing a data
    plane defined by 743 manifests across 12 applications.}
    \label{tab:cp_thinness}
  \end{table}

  The eight distinct defense layers (identity, impersonation, scope
  injection, module gating, surface filtering, response hardening,
  transport, and network) are implemented in approximately 300 lines of
  middleware code. ``Thin'' is not a metaphor; it is measurable.

% ==============================================================================
  \subsection{Q5: Does the boundary reduce operational churn?}
  We evaluate whether governance moves from redeploys to data operations:
  \begin{itemize}
    \item number of deployments triggered by tenant/module enablement changes,

    \item mean time from ``policy edited'' to ``platform reflects change,''

    \item incidence of production hotfixes caused by policy drift across
      services.
  \end{itemize}

  \paragraph{Expected outcome.}
  When the control plane owns tenancy and capability contracts, tenant
  onboarding and module toggles should require fewer code changes and
  fewer redeploy cycles.

  Table~\ref{tab:onboard_ops} shows the six most common tenant lifecycle
  operations and whether each requires a code deployment or only a data
  change.

  \begin{table}[H]
    \centering
    \begin{tabularx}{\textwidth}{l c X}
      \toprule
      \textbf{Operation} & \textbf{Requires Deploy?} & \textbf{Mechanism} \\
      \midrule
      Add new tenant          & No  & DynamoDB record insert \\
      Enable/disable module   & No  & DynamoDB record update (module list) \\
      Change module layout    & No  & DynamoDB \texttt{moduleLayout} JSON field \\
      Add impersonation target & No & Cognito group assignment \\
      Change schema prefix    & No  & DynamoDB tenant record update \\
      Add a \emph{new module type} & Yes & New router + \texttt{@requires\_module} tag \\
      \bottomrule
    \end{tabularx}
    \caption{Tenant lifecycle operations. Five of six common operations
    are data-only changes that take effect without a code deployment,
    reducing operational churn.}
    \label{tab:onboard_ops}
  \end{table}

% ==============================================================================
  \subsection{Q6: How much does surface minimization reduce the visible API?}
  \label{sec:eval_surface}
  The paper claims OpenAPI filtering reduces surface drift. We quantify
  the reduction by counting endpoints visible to tenants with different
  module configurations.

  \begin{table}[H]
    \centering
    \begin{tabularx}{\textwidth}{X c c}
      \toprule
      \textbf{Tenant profile} & \textbf{Visible endpoints} & \textbf{Reduction} \\
      \midrule
      Full platform (all modules)         & 48 & --- \\
      Single module (e.g., Grid only)     & $\sim$16 & $\sim$67\% \\
      Core only (no optional modules)     & 5 & $\sim$90\% \\
      \bottomrule
    \end{tabularx}
    \caption{API surface reduction by OpenAPI filtering. A tenant with a
    single enabled module sees roughly one-third of the total surface;
    a tenant with no optional modules sees only the five core endpoints
    (\texttt{/v2/me}, \texttt{/v1/me}, \texttt{/v1/tenant},
    \texttt{/v1/openapi.json}, walkthrough).}
    \label{tab:surface_reduction}
  \end{table}

  \paragraph{Why this matters.}
  In multi-tenant platforms, surface drift is not merely a documentation
  nuisance; every exposed endpoint is an attack surface and a support
  liability. A 67 to 90\% reduction in visible endpoints for limited-module
  tenants eliminates entire classes of accidental discovery and
  confused-client bugs.

% ==============================================================================
  \subsection{Q7: What does governance concentration save?}
  \label{sec:eval_counterfactual}
  To quantify the value of the control plane boundary, we consider a
  counterfactual: what if each data plane application independently
  implemented its own governance?

  \begin{table}[H]
    \centering
    \begin{tabularx}{\textwidth}{l c c X}
      \toprule
      \textbf{Concern} & \textbf{With CP} & \textbf{Without (per-svc)} & \textbf{Risk avoided} \\
      \midrule
      JWT validation      & 1 impl.\ (73 LoC) & 12 impls.\ & Inconsistent token handling \\
      Tenant lookup       & 1 DynamoDB call     & 12 DB lookups & Stale or missing tenant records \\
      Impersonation logic & 1 allow-list (70 LoC) & 12 scattered checks & Partial or missing allow-lists \\
      Module gating       & 1 decorator          & ad-hoc per svc & Ungated endpoints \\
      Schema prefix       & 1 injection point    & 12 trust-the-client & Confused-deputy risk $\times$12 \\
      OpenAPI filtering   & 1 dynamic filter     & impossible / manual & Permanent surface drift \\
      \bottomrule
    \end{tabularx}
    \caption{Counterfactual analysis: governance with vs.\ without the
    control plane boundary. The control plane replaces an estimated
    $\sim$3{,}600 lines of scattered governance code (300 lines
    $\times$ 12 services) with a single 248-line middleware file.}
    \label{tab:counterfactual}
  \end{table}

  The savings are not merely in lines of code; they are in
  \emph{consistency}. Twelve independent implementations of tenant
  scoping means twelve opportunities for one of them to trust a
  client-provided \texttt{schema} parameter. One implementation means
  one place to audit, one place to patch, and one place to get right.

% ==============================================================================
  \subsection{Q5: Is governance centralized or scattered?}
  Section~\ref{sec:problem} identifies \emph{policy sprawl}, the
  tendency for each service to re-derive tenancy and authorization
  independently, as a root cause of governance drift. This question
  evaluates whether the architecture achieves centralization in
  practice, not merely in diagrams.

  \paragraph{Code evidence.}
  We count the files in the codebase that make governance
  decisions, authentication, tenant resolution, module authorization,
  and scope derivation:
  \begin{itemize}
    \item \textbf{Authentication and context assembly:}
      \texttt{auth\_middleware} in \texttt{middleware.py} (1~file).
      This single middleware resolves the actor identity, validates
      impersonation, and assembles the full effective context.

    \item \textbf{Module-level authorization:}
      the \texttt{@requires\_module} decorator in
      \texttt{auth/decorators.py} (1~file). Every module-specific
      router endpoint (Horizon, Grid, DataMap) is gated by this
      decorator, which checks whether the effective context includes
      the required module; submodule-only endpoints also use
      \texttt{@requires\_submodule}.

    \item \textbf{Scope derivation:}
      \texttt{schema\_prefix} computation in \texttt{middleware.py}
      (same file as authentication). The prefix is derived from the
      tenant record and injected into downstream calls.
  \end{itemize}

  In total, \textbf{three governance mechanisms reside in two files},
  both inside the controller Lambda. Module Lambdas
  (\texttt{horizon\_lambda/}, \texttt{grid\_lambda/}) contain
  \emph{zero authentication or tenant-resolution middleware}. They
  receive a pre-validated scope and execute within it. This is the
  structural counterpart of Figure~\ref{fig:arch_overview}: governance is not
  merely ``recommended to be centralized'', it is architecturally
  impossible for a module Lambda to improvise its own tenancy, because
  the relevant code does not exist there.

% ==============================================================================
  \subsection{Q6: Does context separation support auditable
  impersonation?}
  Design goal G1 requires that actor identity and effective context
  remain structurally separate, especially during impersonation. This
  question evaluates whether the separation is enforced and observable.

  \paragraph{Code evidence.}
  The \texttt{/v2/me} endpoint (routers/user.py) returns a response
  containing distinct \texttt{actor} and \texttt{view\_as} objects.
  These two structures are never merged: the actor reflects the
  authenticated user; \texttt{view\_as} reflects the tenant context
  they are operating in. When no impersonation is active, both point
  to the same identity, but the structural separation is still
  present, consuming code never needs to branch on ``am I
  impersonating?'' to find the right field.

  Before accepting an impersonation request, \texttt{auth\_middleware}
  validates the target tenant against the user's
  \texttt{available\_impersonations} list. If the target is not in the
  list, the request is rejected before any downstream processing
  occurs.

  Impersonation is also visible in telemetry. APM labels include
  \texttt{is\_impersonating}, \texttt{actual\_tenant}, and
  \texttt{viewing\_tenant} (user-context.ts), ensuring that support
  sessions are distinguishable from normal traffic in traces and
  dashboards. Frontend tests (profile.spec.ts) validate the
  actor/\texttt{view\_as} structure, confirming the contract is
  enforced across the stack.

% ==============================================================================
  \subsection{Interpretation}
  The evaluation is not a benchmark contest; it is a correctness and
  operability check. Together, Q1--Q6 cover all five design goals:
  \begin{itemize}
    \item \textbf{G1} (explicit context) is evaluated by Q6: actor and
      effective identities are structurally separate, impersonation is
      validated and visible in telemetry.

    \item \textbf{G2} (server-injected scoping) is evaluated by Q2:
      \texttt{schema\_prefix} is derived server-side and injected into
      every downstream call, with Horizon enforcing its presence.

    \item \textbf{G3} (minimized surface) is evaluated by Q3: OpenAPI
      filtering removes endpoints and schemas per tenant context.

    \item \textbf{G4} (cache-friendly contracts) is evaluated by Q1:
      ETag-based revalidation reduces payload transfer for
      \texttt{/v2/me}.

    \item \textbf{G5} (operational clarity) is evaluated by Q4 and Q5:
      governance changes are data operations (Q4), and policy logic is
      centralized in two files within the controller Lambda (Q5).
  \end{itemize}

  Q5 additionally confirms that governance centralization is structural,
  not merely conventional: module Lambdas contain zero authentication
  or tenant-resolution code. Q6 confirms that context separation is
  maintained across the full stack, from middleware to APM labels to
  frontend tests. If \texttt{/v2/me} is fast, scope mismatches are
  absent, surface drift is reduced, governance is centralized, and
  impersonation is auditable, then the boundary is doing its job:
  keeping governance explicit and keeping compute focused on execution.

  The quantitative results reinforce this: the entire governance layer
  fits in $\sim$6{,}200 lines of Python (Q4), five of six
  common tenant operations require no deployment (Q5), surface filtering
  removes 67 to 90\% of endpoints for limited-module tenants (Q6), and
  the alternative, per-service governance, would scatter $\sim$3{,}600
  lines of policy logic across twelve independent implementations (Q7).
  The boundary is not merely conceptually clean; it is measurably
  economical.
