  \section{Related Work}
  \label{sec:related}
  \paragraph{Control plane vs.\ data plane as a systems pattern.}
  The separation of decision-making from execution is foundational in
  systems architecture. Kubernetes formalizes this distinction inside the
  cluster: control plane components manage desired state while worker
  nodes execute workloads \cite{k8s_components,k8s_architecture}. Managed
  Kubernetes offerings extend the idea by externalizing control plane operations
  behind a service boundary \cite{eks_docs}. This paper applies the
  split to multi-tenant governance: tenancy and scope are decisions;
  compute is execution.

  \paragraph{Multi-tenant Kubernetes frameworks.}
  Prior work has explored explicit control-plane/data-plane isolation for
  multi-tenant Kubernetes, including frameworks that provide per-tenant
  control planes while sharing underlying compute resources (e.g., VirtualCluster)
  \cite{virtualcluster}. Our focus differs: we do not attempt to re-slice
  Kubernetes itself; we keep tenant governance in a platform control plane
  and propagate explicit scope into execution substrates (Kubernetes, ECS/Fargate,
  and module backends).

  \paragraph{Policy engines and authorization languages.}
  OPA provides a general-purpose policy engine that centralizes decision
  logic and evaluation \cite{opa_docs}. Cedar (and Amazon Verified
  Permissions) formalize policy evaluation with a dedicated language and
  service boundary \cite{cedar_docs,aws_verified_permissions}. Our work
  is compatible with these systems but focuses on a narrower objective: producing
  \emph{cacheable governance contracts} (context, surface, scope) that
  downstream systems consume without improvising tenancy.

  \paragraph{Authorization at scale.}
  Zanzibar demonstrates authorization as a global consistency problem and
  highlights the importance of uniform interpretation of identity and context
  across services \cite{zanzibar}. While \system{} does not implement Zanzibar's
  relationship store, the lesson applies directly: once systems scale, "context
  correctness" becomes a distributed property, not a local detail. Our
  design addresses this through explicit contracts and scope injection
  rather than relationship evaluation.

  \paragraph{Serverless APIs for governance.}
  Serverless functions are a natural fit for governance endpoints: spiky
  traffic, small payloads, and a strong need for structured logs
  \cite{aws_lambda_docs,aws_powertools}. The main tradeoff is cold-start
  tail latency; we mitigate this by keeping the control plane thin and
  using HTTP conditional requests for frequent identity contracts
  \cite{rfc9110}.

  \paragraph{Workflow orchestration and batch execution.}
  Airflow is a widely used workflow platform for building and operating batch-oriented
  pipelines \cite{airflow_docs}. Executing tasks in isolated containers
  (e.g., ECS/Fargate) is a pragmatic strategy for burst compute where
  operational simplicity and isolation matter
  \cite{ecs_runtask,fargate_docs}. In our architecture, workflows
  execute jobs, but job scope is derived from authenticated tenant context
  rather than client-supplied parameters.
