 \section{Limitations and Future Work}
  \label{sec:futurework} The control-plane/data-plane split buys clarity,
  but it does not buy immortality. It changes where complexity lives:
  less in scattered services, more in contracts, schemas, and governance
  tooling. This section names the sharp edges and where we intend to
  sand them down.

% ==============================================================================
  \subsection{Contract evolution and versioning}
  Today, \texttt{/v2/me} is a stable contract that the web tier depends
  on heavily. As the platform grows, we must avoid two failure modes:
  \begin{itemize}
    \item \textbf{Contract drift:} clients assume fields that are no
      longer correct.

    \item \textbf{Contract explosion:} every new feature invents a new "me-like"
      endpoint.
  \end{itemize}
  Future work includes explicit contract versioning (beyond path
  versioning), schema documentation, and compatibility tests that run in
  CI against representative tenant/role contexts.

% ==============================================================================
  \subsection{Cache correctness: cheap revalidation without accidental sharing}
  Conditional requests (ETag/\texttt{If-None-Match}) make \texttt{/v2/me}
  cheaper \cite{rfc9110}. However, multi-tenant systems must treat caching
  as a correctness topic, not a performance topic. Future work includes:
  \begin{itemize}
    \item tightening cache directives where shared intermediaries exist,

    \item adding a policy version field so caches can be invalidated deterministically,

    \item and building explicit "cache safety tests" that ensure responses
      are not reusable across principals.
  \end{itemize}

% ==============================================================================
  \subsection{Stronger scope propagation: beyond \texttt{schema\_prefix}}
  Today, \texttt{schema\_prefix} is an effective scoping signal because it
  is derived from authenticated context and injected downstream. The
  limitation is that it is \emph{conventional}: downstream code must
  continue to respect it. Future work includes stronger propagation mechanisms:
  \begin{itemize}
    \item cryptographically signed scope tokens validated by module backends,

    \item and/or a standardized "tenant context envelope" passed through
      every hop (web tier $\rightarrow$ control plane $\rightarrow$
      module $\rightarrow$ database).
  \end{itemize}
  The goal is not to add ceremony; it is to make tenant confusion mechanically
  difficult.

% ==============================================================================
  \subsection{Policy authoring: humans deserve better than raw JSON}
  When policy is data, authoring experience becomes a production
  dependency. Strict parsing is good, but future work should include:
  \begin{itemize}
    \item versioned JSON schemas and offline validation tools,

    \item a review workflow with diffs and previews for tenant/module changes,

    \item and a "staged rollout" mechanism so policy changes can be deployed
      gradually.
  \end{itemize}
  This keeps governance flexible without making it fragile.

% ==============================================================================
  \subsection{OpenAPI filtering: keeping surface minimization provable}
  Filtering OpenAPI reduces surface drift, but it introduces its own
  correctness obligation: the filtered spec must match runtime behavior.
  Future work includes:
  \begin{itemize}
    \item automated spec-vs-runtime consistency checks,

    \item a test harness that enumerates enabled modules and asserts that
      endpoints are both (i) present in the filtered spec and (ii) callable
      when authorized,

    \item and explicit "module tag" governance so tag mistakes cannot leak
      endpoints.
  \end{itemize}

% ==============================================================================
  \subsection{Airflow \& batch execution: toward a first-class job contract}
  Airflow and ECS/Fargate provide a pragmatic batch substrate, but jobs currently
  encode scope as parameters (env/args). Future work includes a first-class
  "job intent" contract emitted by the control plane:
  \begin{itemize}
    \item a structured payload that includes tenant scope, job identity,
      and trace context,

    \item consistent retry semantics and failure classification,

    \item and standardized audit logs that tie job execution back to actor/effective
      context.
  \end{itemize}
  This would make batch execution a citizen of the governance model, not
  a tolerated exception.

% ==============================================================================
  \subsection{Multi-region control plane strategy}
  The platform already runs in multiple regions for resilience (as operational
  reality demands). Future work includes making the control plane explicitly
  region-aware:
  \begin{itemize}
    \item routing strategy for the web tier,

    \item failure-mode playbooks (what happens when a region is
      unhealthy),

    \item and tested procedures for policy/config replication where necessary.
  \end{itemize}

  In summary: the core pattern holds, but the tooling around it must mature.
  A thin control plane stays thin only if we refuse to turn it into a
  dumping ground.

  \bgroup\renewcommand{\addcontentsline}[3]{}\section*{Key Takeaways}\egroup
  \begin{itemize}
    \item \textbf{Make tenancy explicit everywhere.} Treat tenant
      context as a first-class artifact: keep \emph{actor} identity
      immutable, represent \emph{effective} context explicitly, and refuse
      to let downstream systems guess.

    \item \textbf{Decide at the edge, execute in the fabric.} A thin
      serverless control plane is a governance engine: it authenticates,
      derives scope, materializes contracts, and minimizes surface. The container
      data plane should do what it does best: run workloads.

    \item \textbf{Scope injection beats scope suggestion.} Derive
      scoping signals (e.g., \texttt{schema\_prefix}) from authenticated
      context and inject them into delegated calls and jobs. This is the
      practical antidote to confused-deputy failures.

    \item \textbf{Stop advertising what you will deny.} Filter OpenAPI
      and gate routers by enabled modules so clients do not discover
      endpoints that should not exist for their context.

    \item \textbf{Make "who am I?" cheap enough to ask often.} Use ETags
      and conditional requests for high-frequency context contracts. In multi-tenant
      systems, correctness plus caching is not an optimization; it is stability.
  \end{itemize}