% Serverless Control Plane, Container Data Plane.
% Canonical full version for arXiv / technical-report use.

\documentclass[11pt, reqno]{amsart}

% Page and typography.
\usepackage[margin=1in]{geometry}
\usepackage{microtype}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{enumitem}
\usepackage{amsmath, amsfonts, amssymb}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{float}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{listings}
\usepackage{tikz}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\usetikzlibrary{arrows.meta, positioning, fit, calc, backgrounds, shapes.geometric}

\setlength{\emergencystretch}{2em}
\setlist[itemize]{leftmargin=1.6em, itemsep=0.25em, topsep=0.35em}
\setlist[enumerate]{leftmargin=1.8em, itemsep=0.25em, topsep=0.35em}
\raggedbottom

% Colors.
\definecolor{CausifyBlue}{RGB}{3,105,161}
\definecolor{CausifyBg}{RGB}{248,248,248}
\definecolor{CausifyFrame}{RGB}{220,220,220}
\definecolor{CausifyComment}{RGB}{110,110,110}
\definecolor{CausifyKeyword}{RGB}{36,74,140}
\definecolor{CausifyString}{RGB}{24,110,90}
\definecolor{CausifyNumber}{RGB}{125,125,125}

% Listings.
\lstdefinestyle{causify}{
  basicstyle=\ttfamily\footnotesize,
  backgroundcolor=\color{CausifyBg},
  frame=single,
  rulecolor=\color{CausifyFrame},
  framerule=0.4pt,
  framesep=6pt,
  xleftmargin=2pt,
  xrightmargin=2pt,
  numbers=left,
  numberstyle=\tiny\color{CausifyNumber},
  numbersep=8pt,
  keywordstyle=\bfseries\color{CausifyKeyword},
  commentstyle=\itshape\color{CausifyComment},
  stringstyle=\color{CausifyString},
  showstringspaces=false,
  breaklines=true,
  breakatwhitespace=true,
  postbreak=\mbox{\textcolor{CausifyNumber}{$\hookrightarrow$}\space},
  tabsize=2,
  columns=fullflexible,
  keepspaces=true,
  captionpos=b,
  aboveskip=0.8\baselineskip,
  belowskip=0.8\baselineskip
}

\lstdefinestyle{json}{
  style=causify,
  morestring=[b]",
  literate={\\_}{{\_}}1
}

\lstdefinestyle{python}{
  style=causify,
  morekeywords={def,class,return,raise,if,else,elif,for,in,not,and,or,True,False,None,with,as,try,except,finally,import,from,lambda}
}

% Macros.
\newcommand{\system}{\textsc{Causify Platform}}
\newcommand{\console}{\textsc{Causify Console}}
\newcommand{\control}{control plane}
\newcommand{\dataplane}{data plane}
\newcommand{\viewas}{\texttt{view\_as}}
\newcommand{\code}[1]{\texttt{\detokenize{#1}}}
\newcommand{\scopeprefix}{\code{schema_prefix}}

\newenvironment{authorbio}[1]{%
  \par\noindent\textbf{#1}\nobreakspace\ignorespaces
}{\par\vspace{0.4em}}

\title[Serverless Control Plane]{Serverless Control Plane, Container Data Plane\\
\large A Governance Boundary for Multi-Tenant AI Platforms}

\author[Ghasemnezhad et al.]{Shayan Ghasemnezhad, Giacinto Paolo (GP) Saggese, Paul Smith, Heanh Sok, and Vedanshubhai Joshi}
\date{July 2026}

\begin{document}

\begin{abstract}
Multi-tenant AI platforms fail in a plain, dangerous way. Services learn to make their own guesses about identity, tenancy, and capability. A query service trusts a caller-supplied tenant hint. A job runner accepts a schema prefix from a request payload. An OpenAPI document advertises endpoints for modules a tenant never bought. None of these decisions looks dramatic in isolation. Together, they turn governance into folklore.

We describe a production architecture that puts the folklore behind a boundary. A serverless control plane materializes identity, effective tenant context, module capability, API surface, impersonation state, and tenant-scoping signals. A container data plane executes workloads in Kubernetes, Lambda-backed modules, and on-demand Fargate jobs without owning policy. The important property is not AWS Lambda or EKS by themselves; it is the split. The control plane decides and attests. The data plane executes inside injected scope.

We report eight months of operational experience at Causify AI. The system concentrates shared governance in a 135-line Lambda middleware spine, unifies enforcement across 12 data-plane services, reduces visible API surface by 67--90\% for limited-module tenants from a 48-endpoint base, and moves five of six common tenant lifecycle operations to reviewed data changes instead of deployments. From 1,368 production observations, full \code{/v2/me} materialization at 512 MB shows 856 ms P50 latency; ETag revalidation handles 91\% of steady-state traffic at 68 ms P50. Structured-log correlation found zero actor/effective-tenant mismatches within the limits of the detection method. The pattern has been useful in production: make tenancy explicit, inject scope downstream, and stop asking every service to rediscover the truth.
\end{abstract}

\maketitle
\markboth{GHASEMNEZHAD ET AL.}{SERVERLESS CONTROL PLANE}

\noindent\textbf{Keywords.} multi-tenant SaaS; AI platform; authorization; serverless architecture; control plane; data plane; scope injection; confused deputy; API governance; AWS Lambda; Kubernetes

\setcounter{tocdepth}{1}
\tableofcontents

\input{01_introduction.tex}
\input{02_problem_statement.tex}
\input{03_architecture_overview.tex}
\input{04_control_plane_design.tex}
\input{05_container_data_plane.tex}
\input{06_security.tex}
\input{07_operations_reliability.tex}
\input{08_evaluation.tex}
\input{09_related_work.tex}
\input{10_limitations_future_work.tex}
\input{11_conclusion.tex}

\bibliographystyle{plainurl}
\bibliography{references}

\section*{About the Authors}

\begin{authorbio}{Shayan Ghasemnezhad}
is Infrastructure and DevOps Lead at Causify AI, Turin, Italy. His research interests include platform engineering and cloud solutions architecture, production AI/ML systems, and secure, reliable infrastructure operations. Ghasemnezhad received his Bachelor's degree in Digital Management from Ca' Foscari University of Venice. Contact him at shayan@causify.ai.
\end{authorbio}

\begin{authorbio}{Giacinto Paolo (GP) Saggese}
is Co-founder and CTO at Causify AI, Potomac, Maryland, USA. His research interests include machine learning, fintech, and the commercialization of research. Saggese received his Ph.D. degree in Electrical and Computer Engineering from the University of Illinois Urbana-Champaign. He is an Adjunct Professor at the University of Maryland, College Park. Contact him at gp@causify.ai.
\end{authorbio}

\begin{authorbio}{Paul Smith}
is Co-founder and Chief Scientist at Causify AI, Cary, North Carolina, USA. His research interests include causal inference, Bayesian modeling, and time-series prediction. Smith received his Ph.D. degree in Mathematics from UCLA. Contact him at paul@causify.ai.
\end{authorbio}

\begin{authorbio}{Heanh Sok}
is Software Engineer at Causify AI, College Park, Maryland, USA. His research interests include big data systems, distributed systems, and AI/ML engineering and operations. Sok received his Master's degree in Data Science from the University of Maryland, College Park. Contact him at h.sok@causify.ai.
\end{authorbio}

\begin{authorbio}{Vedanshubhai Joshi}
is Software Engineer at Causify AI, Indiana, USA. His research interests include platform infrastructure, multi-tenant API design, and data engineering pipelines. Joshi received his Master's degree in Computer Science from Purdue University. Contact him at v.joshi@causify.ai.
\end{authorbio}
\end{document}
